---
title: "Log regression & regression tree"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
movies <- read.csv(file = "C:/Users/devon/Desktop/Data311/project/IMDB-Movie-Data.csv")
attach(movies)
slr <- lm(Rank ~ Revenue..Millions.) 
summary(slr)
```

## Fixing NA values
varibale named NewMovieDS has no NA values
```{r}

any(is.na(movies))
sum(is.na(movies))
colSums(is.na(movies))
NewMovieDS <- na.omit(movies)
any(is.na(NewMovieDS))
nrow(NewMovieDS)
```

## Logistic regression
For rating
```{r}
simlog<-glm(factor(Rating)~Year+Votes+Metascore,data=NewMovieDS, family=binomial)
#table(predict(simlog, type="response")>0.5, factor(Ratings)
summary(simlog)
summary(simlog)$coef
```
Logistic doesn't look to good at predicting rating (this is after I got rid of the predictors with the worst p values)

For rank
```{r}
simlog2<-glm(factor(Rank)~+Year+Runtime..Minutes.+Votes,data=NewMovieDS, family=binomial)
summary(simlog2)
summary(simlog2)$coef
```
Not to bad for rank. p value of Year increases if runtime is taken out

FOr Genre
```{r}
simlog3<-glm(factor(Genre)~Rank+Rating+Votes+Revenue..Millions.+Metascore,data=NewMovieDS, family=binomial)
summary(simlog3)
summary(simlog3)$coef
```
Logistic regression seems to be fairly good at predicting genre as all p values look to be significatn (except runtime)

For Revenue
```{r}
simlog4<-glm(factor(Revenue..Millions.)~Rating+Runtime..Minutes.+Votes,data=NewMovieDS, family=binomial)
summary(simlog4)
summary(simlog4)$coef
```
Log regression not bad for revenue but not good either (pvalue for votes increases if rating taken out)

For Runtime
```{r}
simlog5<-glm(factor(Runtime..Minutes.)~Rank+Rating+Votes+Revenue..Millions.+Metascore,data=NewMovieDS, family=binomial)
summary(simlog5)
summary(simlog5)$coef
```
At a glance can tell not good for predicting runtime

For votes
```{r}
simlog6<-glm(factor(Votes)~Rank+Rating+Runtime..Minutes.+Revenue..Millions.,data=NewMovieDS, family=binomial)
summary(simlog6)
summary(simlog6)$coef
```
As before at a glance can tell not very good at predicting votes

For MetaScore
```{r}
simlog7<-glm(factor(Metascore)~Rank+Runtime..Minutes.+Rating+Votes,data=NewMovieDS, family=binomial)
summary(simlog7)
summary(simlog7)$coef
```

