---
title: "Log regression & regression tree"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
movies <- read.csv(file = "C:/Users/devon/Desktop/Data311/project/IMDB-Movie-Data.csv")
attach(movies)
slr <- lm(Rank ~ Revenue..Millions.) 
summary(slr)
```

## Fixing NA values
varibale named NewMovieDS has no NA values
```{r}
set.seed(1546)
any(is.na(movies))
sum(is.na(movies))
colSums(is.na(movies))
NewMovieDS <- na.omit(movies)
any(is.na(NewMovieDS))
nrow(NewMovieDS)

train<-sample (1: nrow(NewMovieDS), nrow(NewMovieDS)/2)

movtest <- NewMovieDS[-train, ]
```

## Logistic regression
For rating
```{r}
simlog<-glm(factor(Rating)~Year+Votes+Metascore,data=NewMovieDS, family=binomial)
#table(predict(simlog, type="response")>0.5, factor(Ratings)
summary(simlog)
summary(simlog)$coef

logTest<-glm(factor(Rating)~Year+Votes+Metascore,data=NewMovieDS, family=binomial, subset=train)
logPred <-predict(logTest,factor(movtest$Rating), type="response")
#table(logPred>0.5, factor(movtest$Ratings))
movTest = NewMovieDS[-train, "Rating"]
mean(logPred==movTest)
```
Logistic doesn't look to good at predicting rating (this is after I got rid of the predictors with the worst p values)

For rank
```{r}
simlog2<-glm(factor(Rank)~+Year+Runtime..Minutes.+Votes,data=NewMovieDS, family=binomial)
summary(simlog2)
summary(simlog2)$coef
logTest2<-glm(factor(Rank)~+Year+Runtime..Minutes.+Votes,data=NewMovieDS, family=binomial, subset=train)
logPred2<-predict(logTest2,factor(movtest$Rank), type="response")
#table(logPred>0.5, factor(movtest$Ratings))
movTest2 = NewMovieDS[-train, "Rank"]
mean(logPred2==movTest2)
```
Not to bad for rank. p value of Year increases if runtime is taken out

FOr Genre
```{r}
simlog3<-glm(factor(Genre)~Rank+Rating+Votes+Revenue..Millions.+Metascore,data=NewMovieDS, family=binomial)
summary(simlog3)
summary(simlog3)$coef

logTest3<-glm(factor(Genre)~Rank+Rating+Votes+Revenue..Millions.+Metascore,data=NewMovieDS, family=binomial, subset=train)
logPred3<-predict(simlog3, type="response")
#table(logPred3>0.5, Genre)
movTest3 = NewMovieDS[-train, "Genre"]
mean(logPred3==movTest3)
```
Logistic regression seems to be fairly good at predicting genre as all p values look to be significatn (except runtime)

For Revenue
```{r}
simlog4<-glm(factor(Revenue..Millions.)~Rating+Runtime..Minutes.+Votes,data=NewMovieDS, family=binomial)
summary(simlog4)
summary(simlog4)$coef
```
Log regression not bad for revenue but not good either (pvalue for votes increases if rating taken out)

For Runtime
```{r}
simlog5<-glm(factor(Runtime..Minutes.)~Rank+Rating+Votes+Revenue..Millions.+Metascore,data=NewMovieDS, family=binomial)
summary(simlog5)
summary(simlog5)$coef
```
At a glance can tell not good for predicting runtime

For votes
```{r}
simlog6<-glm(factor(Votes)~Rank+Rating+Runtime..Minutes.+Revenue..Millions.,data=NewMovieDS, family=binomial)
summary(simlog6)
summary(simlog6)$coef
```
As before at a glance can tell not very good at predicting votes

For MetaScore
```{r}
simlog7<-glm(factor(Metascore)~Rank+Runtime..Minutes.+Rating+Votes,data=NewMovieDS, family=binomial)
summary(simlog7)
summary(simlog7)$coef
```

##Regression tree

```{r}
library(tree)
library(MASS)


treeMove<- tree(NewMovieDS$Rating~Rank+Genre+Year+Runtime..Minutes.+Votes+Revenue..Millions.+Metascore,NewMovieDS)
summary(treeMove)
plot(treeMove)
text(treeMove, pretty=0)

cvMove<- cv.tree(treeMove)
summary(cvMove)
plot(cvMove)

pruneMove <- prune.tree(treeMove, best =8)
plot(pruneMove)
text(pruneMove, pretty=0)



treeMove2<- tree(NewMovieDS$Rating~Rank+Genre+Year+Runtime..Minutes.+Votes+Revenue..Millions.+Metascore,NewMovieDS, subset=train)
summary(treeMove2)
plot(treeMove2)
text(treeMove2, pretty=0)

yhat<- predict(treeMove2, newdata=movtest)
MoveTest<- NewMovieDS[-train, "Rating"]

mean((yhat-MoveTest)^2)
```
MSE is 0.4491893 so this tree predicts rating within .67 of the true response. Which is not bad.


For Rank

```{r}
treeMove3<- tree(NewMovieDS$Rank~Rating+Genre+Year+Runtime..Minutes.+Votes+Revenue..Millions.+Metascore,NewMovieDS)
summary(treeMove3)
plot(treeMove3)
text(treeMove3, pretty=0)

cvMove2<- cv.tree(treeMove3)
summary(cvMove2)
plot(cvMove2)

pruneMove2 <- prune.tree(treeMove3, best =6)
plot(pruneMove2)
text(pruneMove2, pretty=0)



treeMove4<- tree(NewMovieDS$Rank~Rating+Genre+Year+Runtime..Minutes.+Votes+Revenue..Millions.+Metascore,NewMovieDS, subset=train)
summary(treeMove4)
plot(treeMove4)
text(treeMove4, pretty=0)

yhat2<- predict(treeMove4, newdata=movtest)
MoveTest2<- NewMovieDS[-train, "Rank"]

mean((yhat2-MoveTest2)^2)
```
MSE is 56096.81 so this tree predicts rank within 236.85 of the true response. Not very good


For genre
```{r}
treeMove5<- tree(NewMovieDS$Genre~Rank+Year+Runtime..Minutes.+Rating+Votes+Revenue..Millions.+Metascore,NewMovieDS)
summary(treeMove5)
plot(treeMove5)
text(treeMove5, pretty=0)

cvMove3<- cv.tree(treeMove5)
summary(cvMove3)
plot(cvMove3)

pruneMove3 <- prune.tree(treeMove5, best =6)
plot(pruneMove3)
text(pruneMove3, pretty=0)



treeMove6<- tree(NewMovieDS$Genre~Rank+Year+Runtime..Minutes.+Rating+Votes+Revenue..Millions.+Metascore,NewMovieDS, subset=train)
summary(treeMove6)
plot(treeMove6)
text(treeMove6, pretty=0)

yhat3<- predict(treeMove6, newdata=movtest)
MoveTest3<- NewMovieDS[-train, "Genre"]

mean((yhat3-MoveTest3)^2)
```
MSE is NA which makes sense as genre is catagorical so a regression tree is not a good method of prediction


FOr revenue
```{r}
treeMove7<- tree(NewMovieDS$Revenue..Millions.~Rank+Genre+Year+Runtime..Minutes.+Rating+Votes+Metascore,NewMovieDS)
summary(treeMove7)
plot(treeMove7)
text(treeMove7, pretty=0)

cvMove4<- cv.tree(treeMove7)
summary(cvMove4)
plot(cvMove4)

pruneMove4 <- prune.tree(treeMove7, best =7)
plot(pruneMove4)
text(pruneMove4, pretty=0)



treeMove8<- tree(NewMovieDS$Revenue..Millions.~Rank+Genre+Year+Runtime..Minutes.+Rating+Votes+Metascore,NewMovieDS, subset=train)
summary(treeMove8)
plot(treeMove8)
text(treeMove8, pretty=0)

yhat4<- predict(treeMove8, newdata=movtest)
MoveTest4<- NewMovieDS[-train, "Revenue..Millions."]

mean((yhat4-MoveTest4)^2)
```
MSE for revenue is 7376.818 so this tree predicts revenue within 85.88 of the true response.This is not very good but given the large variation of the revenue data its hard to tell
if this is bad.


For runtime
```{r}
treeMove9<- tree(NewMovieDS$Runtime..Minutes.~Rank+Genre+Year+Rating+Votes+Revenue..Millions.+Metascore,NewMovieDS)
summary(treeMove9)
plot(treeMove9)
text(treeMove9, pretty=0)

cvMove5<- cv.tree(treeMove9)
summary(cvMove5)
plot(cvMove5)

pruneMove5 <- prune.tree(treeMove9, best =7)
plot(pruneMove5)
text(pruneMove5, pretty=0)



treeMove10<- tree(NewMovieDS$Runtime..Minutes.~Rank+Genre+Year+Rating+Votes+Revenue..Millions.+Metascore,NewMovieDS, subset=train)
summary(treeMove10)
plot(treeMove10)
text(treeMove10, pretty=0)

yhat5<- predict(treeMove10, newdata=movtest)
MoveTest5<- NewMovieDS[-train, "Runtime..Minutes."]

mean((yhat5-MoveTest5)^2)
```
MSE for runtime is 270.74 so this tree predicts runtime to within 16.45 minutes. This is not a bad prediction of the runtime


for votes
```{r}
treeMove11<- tree(NewMovieDS$Votes~Rank+Genre+Year+Runtime..Minutes.+Rating+Revenue..Millions.+Metascore,NewMovieDS)
summary(treeMove11)
plot(treeMove11)
text(treeMove11, pretty=0)

cvMove6<- cv.tree(treeMove11)
summary(cvMove6)
plot(cvMove6)

pruneMove6 <- prune.tree(treeMove11, best =7)
plot(pruneMove6)
text(pruneMove6, pretty=0)



treeMove12<- tree(NewMovieDS$Votes~Rank+Genre+Year+Runtime..Minutes.+Rating+Revenue..Millions.+Metascore,NewMovieDS, subset=train)
summary(treeMove12)
plot(treeMove12)
text(treeMove12, pretty=0)

yhat6<- predict(treeMove12, newdata=movtest)
MoveTest6<- NewMovieDS[-train, "Votes"]

mean((yhat6-MoveTest6)^2)
```
MSE for votes is 16857229085 so this tree predicts votes to within 129835.39 votes. Even with the large variance in votes data this prediction is quite bad and shows that a regression tree likely does not work for this response variable


for metascore
```{r}
treeMove13<- tree(NewMovieDS$Metascore~Rank+Genre+Year+Runtime..Minutes.+Rating+Votes+Revenue..Millions.,NewMovieDS)
summary(treeMove13)
plot(treeMove13)
text(treeMove13, pretty=0)

cvMove7<- cv.tree(treeMove13)
summary(cvMove7)
plot(cvMove7)

pruneMove7 <- prune.tree(treeMove13, best =4)
plot(pruneMove7)
text(pruneMove7, pretty=0)



treeMove14<- tree(NewMovieDS$Metascore~Rank+Genre+Year+Runtime..Minutes.+Rating+Votes+Revenue..Millions.,NewMovieDS, subset=train)
summary(treeMove14)
plot(treeMove14)
text(treeMove14, pretty=0)

yhat7<- predict(treeMove14, newdata=movtest)
MoveTest7<- NewMovieDS[-train, "Metascore"]

mean((yhat7-MoveTest7)^2)
```
MSE for metascore is 174.9464 so this tree predicts metascore within 13.23 points. While this might not be the best model for predicting metascore it is not bad.


##Boosting
For rating
```{r}
library(gbm)
library(gclus)
rain<-sample (1: nrow(NewMovieDS), nrow(NewMovieDS)/2)
lMovie <-NewMovieDS[rain,]
move <- NewMovieDS[-train, ]
moveBoost<-gbm(Rating~Rank+Genre+Year+Runtime..Minutes.+Votes+Revenue..Millions.+Metascore,distribution="gaussian", data=lMovie, n.trees=5000, interaction.depth=1)
summary(moveBoost)
table(lMovie$Rating, predict(moveBoost, newdata=lMovie, type="response", n.trees=5000)>0.5 )


yhatBoost<-predict(moveBoost, newdata=movtest,n.trees=5000)
boostMoveTest<- NewMovieDS[-train, "Rating"]
mean((yhatBoost-boostMoveTest)^2)
```
MSE = 0.2390691  standard deviation = 0.489

For Rank
```{r}
moveBoost2<-gbm(Rank~Genre+Year+Runtime..Minutes.+Rating+Votes+Revenue..Millions.+Metascore,distribution="gaussian", data=lMovie, n.trees=5000, interaction.depth=1)
summary(moveBoost2)
table(lMovie$Rank, predict(moveBoost2, newdata=lMovie, type="response", n.trees=5000)>0.5 )


yhatBoost2<-predict(moveBoost2, newdata=movtest,n.trees=5000)
boostMoveTest2<- NewMovieDS[-train, "Rank"]
mean((yhatBoost2-boostMoveTest2)^2)
```
MSE = 40493.11  SD= 201.23

For Genre
```{r}
moveBoost3<-gbm(Genre~Rank+Year+Runtime..Minutes.+Rating+Votes+Revenue..Millions.+Metascore,distribution="gaussian", data=lMovie, n.trees=5000, interaction.depth=1)
summary(moveBoost3)
table(lMovie$Genre, predict(moveBoost3, newdata=lMovie, type="response", n.trees=5000)>0.5 )


yhatBoost3<-predict(moveBoost3, newdata=movtest,n.trees=5000)
boostMoveTest3<- NewMovieDS[-train, "Genre"]
mean((yhatBoost3-boostMoveTest3)^2)
```
MSE = NA makes semse as categorical not numeric

For Revenue
```{r}
moveBoost4<-gbm(Revenue..Millions.~Rank+Genre+Year+Runtime..Minutes.+Rating+Votes+Metascore,distribution="gaussian", data=lMovie, n.trees=5000, interaction.depth=1)
summary(moveBoost4)
table(lMovie$Revenue..Millions., predict(moveBoost4, newdata=lMovie, type="response", n.trees=5000)>0.5 )


yhatBoost4<-predict(moveBoost4, newdata=movtest,n.trees=5000)
boostMoveTest4<- NewMovieDS[-train, "Revenue..Millions."]
mean((yhatBoost4-boostMoveTest4)^2)
```
MSE= 2760.968  SD=52.54

For runttime
```{r}
moveBoost5<-gbm(Runtime..Minutes.~Rank+Genre+Year+Rating+Votes+Revenue..Millions.+Metascore,distribution="gaussian", data=lMovie, n.trees=5000, interaction.depth=1)
summary(moveBoost5)
table(lMovie$Runtime..Minutes, predict(moveBoost5, newdata=lMovie, type="response", n.trees=5000)>0.5 )


yhatBoost5<-predict(moveBoost5, newdata=movtest,n.trees=5000)
boostMoveTest5<- NewMovieDS[-train, "Runtime..Minutes."]
mean((yhatBoost5-boostMoveTest5)^2)
```
MSE=203.7398  SD=14.27

For votes
```{r}
moveBoost6<-gbm(Votes~Rank+Genre+Year+Runtime..Minutes.+Rating+Revenue..Millions.+Metascore,distribution="gaussian", data=lMovie, n.trees=5000, interaction.depth=1)
summary(moveBoost6)
table(lMovie$Votes, predict(moveBoost6, newdata=lMovie, type="response", n.trees=5000)>0.5 )


yhatBoost6<-predict(moveBoost6, newdata=movtest,n.trees=5000)
boostMoveTest6<- NewMovieDS[-train, "Votes"]
mean((yhatBoost6-boostMoveTest6)^2)
```
MSE=8179454600  SD=90440.34

For metascore
```{r}
moveBoost7<-gbm(Metascore~Rank+Genre+Year+Runtime..Minutes.+Rating+Votes+Revenue..Millions.,distribution="gaussian", data=lMovie, n.trees=5000, interaction.depth=1)
summary(moveBoost7)
table(lMovie$Metascore, predict(moveBoost7, newdata=lMovie, type="response", n.trees=5000)>0.5 )


yhatBoost7<-predict(moveBoost7, newdata=movtest,n.trees=5000)
boostMoveTest7<- NewMovieDS[-train, "Metascore"]
mean((yhatBoost7-boostMoveTest7)^2)
```
MSE=123.7931 SD=11.13
